### 2025-03-28 ms/response_letter.qmd

> 2) Abstract is informative and it reflects the body of the paper.

We appreciate your positive feedback on the abstract.

> 3) The objectives must be modified for a better understanding of the reader.

Thank you for your suggestion.
We have revised the Introduction to clearly state the main objective of the study before presenting the specific questions (line 105-107).

@@

> 4) Line 366-369 & Table 2: In the sentence "In addition, the estimated exponent aligns closely with MST predictions for crown radius, supporting the selection of the power-law model for angiosperms, where the community-level exponent b is near the expected *2/3* ($b_{ang}$=0.625775 0.6168636,0.6345041, Table 2)" and the parameter estimates in Table 2, the values are not consistently presented. Check whole of the manuscript for a uniform presentation of the values.

We have standardized the presentation of all parameter estimates throughout the manuscript, including Table 2 and Tables S6-8 and corresponding text.

> 5) Table 1: The table does not clearly explain how much of a difference in ΔLOOIC is considered significant between the two models.

We have revised the caption of Table 1 to clarify that a $\Delta$LOOIC greater than 4 suggests a meaningful difference in predictive performance (line 238-239).

@@

> 8) Supplementary Table S6 is very congested and hard to read. Please revise for better understanding.

We have reformatted Supplementary Tables S6, S7, and S8 for improved readability and included detailed captions to aid interpretation.

> 9) A revision of the English is needed as some sentences aren't properly structured, and it's often difficult to grasp what the authors are saying.

We have revised the manuscript for language, grammar, and sentence structure to enhance clarity and improve overall readability.
For example, we revised lines 432-438, 448-452 in the Discussion section to ensure sentence brevity and clarity.
Additionally, we updated the figure legends and table captions for greater clarity and carefully cross-checked all technical terms to maintain consistency throughout the text.

@@

> Reviewer #5: Congratulations on a well-written and clear manuscript. Your approach to exploring allometric relationships between DBH, height, wood density, biomass, and crown radius is highly interesting and contributes significantly to the field. While I primarily focus on the data analysis section and found no major concerns, I would like to suggest a few minor corrections and discussion points.

Thank you so much for your encouraging and constructive feedback.

> 1) In Figure 3, the distinction between species with light and dense wood density is presented. However, the criteria used for this classification are not mentioned in the Materials and Methods section. Clarifying this would improve the transparency of your methodology.

Thank you for pointing this out.
We have added a description in the Methods section clarifying the classification criteria (line 178-183).

@@

> 2) Figure 4 appears to have a formatting issue, as the letters 'a' and 'd' are not displayed.

We were unable to see the absence of the letters "a" and "d" in Figure 4 of the generated PDF file.
Nevertheless, we slightly increased the left margin of the panel labels in case clipping was the cause.
We hope this adjustment addresses the concern and would be happy to make further modifications if needed.

> 3) It may be valuable to discuss the potential confounding effects of environmental factors on observed allometric correlations when considering wood density. Given intraspecific variation, wood density may not have a strong impact on those allometric relationships, since Fajardo et al. (2022) suggest that wood density is more closely linked to vessel anatomy than to tree architecture within the same species across different environments.

We agree with your point and have expanded the Discussion section to address potential environmental confounding effects.
We also cited Fajardo et al. (2022) to highlight less clear linkages between tree architecture and intraspecific variation in wood density  (line 392-398).

### 2025-03-28 ms/response_letter.tex

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{3}
\tightlist
\item
  Line 366-369 \& Table 2: In the sentence ``In addition, the estimated
  exponent aligns closely with MST predictions for crown radius,
  supporting the selection of the power-law model for angiosperms, where
  the community-level exponent b is near the expected \emph{2/3}
  (\(b_{ang}\)=0.625775 0.6168636,0.6345041, Table 2)'' and the
  parameter estimates in Table 2, the values are not consistently
  presented. Check whole of the manuscript for a uniform presentation of
  the values.
\end{enumerate}
\end{quote}

We have standardized the presentation of all parameter estimates
throughout the manuscript, including Table 2 and Tables S6-8 and
corresponding text.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{4}
\tightlist
\item
  Table 1: The table does not clearly explain how much of a difference
  in ΔLOOIC is considered significant between the two models.
\end{enumerate}
\end{quote}

We have revised the caption of Table 1 to clarify that a \(\Delta\)LOOIC
greater than 4 suggests a meaningful difference in predictive
performance (line 238-239).

@@

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{1}
\tightlist
\item
  Figure 4 appears to have a formatting issue, as the letters `a' and
  `d' are not displayed.
\end{enumerate}
\end{quote}

We were unable to see the absence of the letters ``a'' and ``d'' in
Figure 4 of the generated PDF file. Nevertheless, we slightly increased
the left margin of the panel labels in case clipping was the cause. We
hope this adjustment addresses the concern and would be happy to make
further modifications if needed.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{2}
\tightlist
\item
  It may be valuable to discuss the potential confounding effects of
  environmental factors on observed allometric correlations when
  considering wood density. Given intraspecific variation, wood density
  may not have a strong impact on those allometric relationships, since
  Fajardo et al.~(2022) suggest that wood density is more closely linked
  to vessel anatomy than to tree architecture within the same species
  across different environments.
\end{enumerate}
\end{quote}

We agree with your point and have expanded the Discussion section to
address potential environmental confounding effects. We also cited
Fajardo et al.~(2022) to highlight less clear linkages between tree
architecture and intraspecific variation in wood density (line 392-398).

### 2025-07-03 ms/ms.qmd
updated ms

^1^ CAS Key Laboratory of Tropical Forest Ecology, Xishuangbanna Tropical Botanical Garden, Chinese Academy of Sciences, Menglun, Yunnan 666303, China

@@

where $f(\cdot)$ predicts the expected natural log tree height or crown radius from observed DBH (and *vice versa*), $\pmb{\beta}_{j}=\{a_j, b_j, k_j\}$ denotes the vector of species-specific parameters, and $\sigma_y$ represents the standard deviation of the observation error.
We considered three functional forms for the relationship between tree allometric variables [@Sullivan2018; @MartinezCano2019]:

@@

We applied the power-law model to estimate tree height and crown radius from DBH, as well as DBH from tree height and/or crown radius.
Specifically, when predicting DBH from height and/or crown radius, we tested four formulations:
(1) using CR and H as separate factors, $\mathrm{DBH} = a_1 \mathrm{CR}^{b_1} \mathrm{H}^{c_1}$, denoted as $CR,H$;
(2) using CR and H as a compound term, $\mathrm{DBH} = a_2 (\mathrm{CR} \times \mathrm{H})^{b_2}$, denoted as $CR \times H$;
(3) using only height, $\mathrm{DBH} = a_3 \mathrm{H}^{b_3}$; and
(4) using only crown radius, $\mathrm{DBH} = a_4 \mathrm{CR}^{b_4}$.
In contrast, the gMM and Weibull models were only used to estimate tree height and crown radius from DBH.
Because these two models specifically describe saturating relationships (@eq-gMM; @eq-wb), their formulation does not allow simply switching the roles of explanatory and response variables.

To compare the performance of different models for the same response variable, we used Pareto-smoothed importance sampling leave-one-out cross-validation (PSIS-LOO; @Vehtari2017).
PSIS-LOO is an accurate and reliable approximation to standard leave-one-out cross-validation (LOO), which is a robust method for comparing models with different numbers of parameters [@Vehtari2017].
We utilized PSIS-LOO to calculate the LOO Information Criterion (LOOIC) for each model.
A model with lower LOOIC is considered to have better predictive accuracy [@Vehtari2017].

@@

where $\gamma_{0}$ and $\gamma_{1}$ are the species-level intercept and slope, respectively, and $\sigma_{\beta}$ is the standard deviation of the species-specific parameters.
To facilitate interpretation and visualization, species were then classified into "light" and "dense" wood groups based on the 10th and 90th percentiles of wood density distribution in the observed dataset, respectively.
This classification allowed us to illustrate the relationships between wood density and the parameters of the best-performing models, enabling an evaluation of how wood density contributes to variation in tree architecture.
Additionally, we explored a model that accounts for covariance between species-specific parameters using a multivariate normal distribution.
However, due to the complexity of this approach, the model did not converge for the nonlinear models, and thus, we did not include it in the final analysis.

Posterior distributions of all parameters were estimated using the No-U-Turn Sampler (NUTS), an adaptive variant of the Hamiltonian Monte Carlo (HMC), algorithm implemented in Stan [@Carpenter2017] using weakly-informative priors [@Gelman2008].
The HMC algorithm uses gradient information to propose new states in the Markov chain, leading to a more efficient exploration of the target distribution than traditional Markov chain Monte Carlo (MCMC) methods that rely on random proposals [@Carpenter2017].
This efficiency allows us to achieve convergence with fewer iterations than traditional MCMC methods.
Inferences were based on 1,000 posterior samples following 1,000 warmup iterations for three parallel chains.
Convergence was assessed using the Gelman–Rubin statistic with a convergence threshold of 1.1 [@Gelman2013] and ensuring effective sample sizes greater than 300 [@Vehtari2021].
We also reported the median and 95\% Bayesian credible intervals (CIs) for each parameter (Supplementary Tables S3-S8).
Estimates were considered significant when 95\% CIs did not overlap with zero.

@@

In CR-DBH allometry, angiosperms exhibited a positive relationship between the scaling factor $a$ and wood density and a negative association between the exponent parameter  $b$ and wood density (@fig-wood1 g, h; @fig-wood2 b).
In gymnosperms, we found a positive association between wood density and the exponent parameter $b$, whereas neither the asymptote parameter $a$ nor the scaling constant $k$ showed significant relationship with this functional trait (@fig-wood1 i, j, k; @fig-wood2 c).

@@

AGB estimates based on the Weibull model ($\text{AGB}_\text{H-WB}$) showed the highest predictive accuracy based both on the lowest root mean square error (RMSE) of `r sprintf("%.3g", data_metrics$metrics_agb_ang$rmse_wb_h)` Mg and the smallest bias of `r sprintf("%.3g", data_metrics$metrics_agb_ang$bias_wb_h * 100)`% (@fig-agb a).
Estimates based on the power-law model ($\text{AGB}_\text{H-PL}$) also showed competitive performance but had higher RMSE and bias compared to the Weibull-based approach, with RMSE of `r sprintf("%.3g", data_metrics$metrics_agb_ang$rmse_pl_h)` Mg and bias of `r sprintf("%.3g", data_metrics$metrics_agb_ang$bias_pl_h * 100)`% (@fig-agb b).
In contrast, models that used estimated DBH performed substantially worse (@fig-agb c-f).

@@

Our result supported a saturating H-DBH relationship in both angiosperm and gymnosperm species, consistent with theoretical expectations [@Falster2003; @Niklas2007] and empirical observations [e.g., @Feldpausch2011; @MartinezCano2019; @Song2024].
A recent study using Tallo suggested the Weibull function as the best model for six plant functional types (PFTs) [@Song2024].
While PFTs-based approaches can illuminate large-scale patterns, they may obscure species-specific variation in allometric scaling, and locally abundant species may bias the overall estimates.
Our best predictive model effectively captures species-specific patterns (Supplementary Table S6) despite underestimating the height of individuals with DBH > 30 cm due to substantial interspecific variation in H–DBH allometric scaling (@fig-compare).
This highlights the advantages of a multilevel modeling approach in addressing systematic differences across species.
Additionally, the metabolic scaling theory (MST) posits logarithmic scaling relationships among different axes of tree size, with an exponent close to 2/3 ($H \propto DBH^{2/3}$, $CR \propto DBH^{2/3}$), based on assumptions regarding mechanical and hydraulic constraints to plant growth [@West1999].
However, our findings revealed that the power-law model exponent of angiosperms (*b* = 0.538) was significantly smaller, while the actual exponent in gymnosperms (*b* = 0.704) was larger than the theoretical value of 2/3 (Supplementary Table S3).
This divergence highlights the importance of separately modeling the allometric relationships of each functional group to accurately capture their distinct scaling patterns.
Moreover, the power-law function performed worse than the saturating functions (Table 1), indicating that the actual H-DBH relationship does not follow the simple scaling laws proposed by MST.

Our analysis also revealed divergent patterns in CR-DBH allometry, highlighting fundamental differences in growth strategies and ecological adaptations between angiosperms and gymnosperms.
In angiosperms, the CR–DBH relationship follows a power-law form, suggesting a capacity for plastic crown expansion in competitive environments [e.g., @Poorter2012a; @Lines2012; @Hulshof2015], a conclusion consistent with a study in tropical moist forests by @MartinezCano2019.
In addition, the estimated exponent aligns closely with MST predictions for crown radius, supporting the selection of the power-law model for angiosperms, where the community-level exponent *b* is near the expected 2/3
($b_{\text{ang}}$ = `r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Median)`
[`r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Lower)`,
`r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Upper)`], Table 2).
In contrast, the CR-DBH allometry in gymnosperms showed an asymptotic relationship best described by the gMM function, with the Weibull function also showing competitive predictive performance (Table 1).
This asymptotic pattern likely arises from structural constraints in gymnosperms growth forms that favour vertical over lateral expansion [e.g., @Lines2012; @Hulshof2015; @Jucker2024], an adaptation that can mitigate risks such as branch breakage under heavy snow loads [e.g., @Mikola1938; @Loehle2016; @Tavankar2019].

@@

Our analysis indicated distinct patterns in which wood density has a greater influence on the interspecific variation in angiosperms than gymnosperms.
Variation in wood density has consequential implications for the mechanical properties of wood and hence influences the mechanical stability of stems and branches [@Sterck1998], ultimately determining the architecture of trees [e.g., @Poorter2006].
Angiosperms have denser wood and a wider wood density range than gymnosperms (see Supplementary Table S2), resulting in greater and more flexible mechanical support [e.g., @Gelder2006; @Swenson2007; @Mo2024].
This allows them to develop a more diverse height and crown to optimize light capture, particularly in environments where light competition is intense, such as tropical regions.
In contrast, gymnosperms predominantly live and often dominate in extreme environments such as boreal and high-elevation regions [@Swenson2007; @Mo2024], prioritising survival and longevity [@Fossdal2024].
Due to their simpler and more homogeneous wood anatomy, these species display lighter wood and a narrower range of wood density [e.g., @Mo2024], which may result in less flexibility in the crown architecture.
These differences underscore how the contrasting evolutionary strategies of angiosperms and gymnosperms shape their architectural diversity, reflecting adaptations to distinct ecological niches.

>>>>

^1^ Yunnan Key Laboratory of Forest Ecosystem Stability and Global Change Response, Xishuangbanna Tropical Botanical Garden, Chinese Academy of Sciences, Menglun, Yunnan 666303, China

@@

where $f(\cdot)$ predicts the expected natural log tree height or crown radius from observed DBH (and *vice versa*), $\pmb{\beta}_{j}=\{a_j, b_j, c_j, k_j\}$ denotes the vector of species-specific parameters, and $\sigma_y$ represents the standard deviation of the observation error.
We considered three functional forms for the relationship between tree allometric variables [@Sullivan2018; @MartinezCano2019]:

@@

We applied the power-law model to estimate tree height and crown radius from DBH, as well as DBH from tree height and/or crown radius.
Specifically, when predicting DBH from height and/or crown radius, we tested four formulations:
(1) using CR and H as separate factors, $\mathrm{DBH} = a_1 \mathrm{CR}^{b_1} \mathrm{H}^{c_1}$, denoted as $CR,H$;
(2) using CR and H as a compound term, $\mathrm{DBH} = a_2 (\mathrm{CR} \times \mathrm{H})^{b_2}$, denoted as $CR \times H$;
(3) using only height, $\mathrm{DBH} = a_3 \mathrm{H}^{b_3}$; and
(4) using only crown radius, $\mathrm{DBH} = a_4 \mathrm{CR}^{b_4}$.
In contrast, the gMM and Weibull models were only used to estimate tree height and crown radius from DBH.
Because these two models specifically describe saturating relationships (@eq-gMM; @eq-wb), their formulation does not allow for simply switching the roles of explanatory and response variables.

To compare the performance of different models for the same response variable, we used Pareto-smoothed importance sampling leave-one-out cross-validation (PSIS-LOO; @Vehtari2017).
PSIS-LOO is an accurate and reliable approximation to standard leave-one-out cross-validation (LOO), a robust method for comparing models with different numbers of parameters [@Vehtari2017].
We utilized PSIS-LOO to calculate the LOO Information Criterion (LOOIC) for each model.
A model with a lower LOOIC is considered to have better predictive accuracy [@Vehtari2017].

@@

where $\gamma_{0}$ and $\gamma_{1}$ are the global-level hyperparameters for intercept and slope, respectively, and $\sigma_{\beta}$ is the standard deviation of the species-specific parameters.
To facilitate interpretation and visualization, species were then classified into "light" and "dense" wood groups based on the 10th and 90th percentiles of wood density distribution in the observed dataset, respectively.
This classification allowed us to illustrate the relationships between wood density and the parameters of the best-performing models, enabling an evaluation of how wood density contributes to variation in tree architecture.
Additionally, we explored a model that accounts for covariance between species-specific parameters using a multivariate normal distribution.
However, due to the complexity of this approach, the model did not converge for the nonlinear models, and thus, we did not include it in the final analysis.

Posterior distributions of all parameters were estimated using the No-U-Turn Sampler (NUTS), an adaptive variant of the Hamiltonian Monte Carlo (HMC), algorithm implemented in Stan [@Carpenter2017] using weakly-informative priors [@Gelman2008].
The HMC algorithm uses gradient information to propose new states in the Markov chain, leading to a more efficient exploration of the target distribution than traditional Markov chain Monte Carlo (MCMC) methods that rely on random proposals [@Carpenter2017].
This efficiency allows us to achieve convergence with fewer iterations than traditional MCMC methods.
Inferences were based on 1,000 posterior samples following 1,000 warmup iterations for three parallel chains.
Convergence was assessed using the Gelman–Rubin statistic with a convergence threshold of 1.1 [@Gelman2013] and ensuring effective sample sizes greater than 300 [@Vehtari2021].
We also reported the median and 95\% Bayesian credible intervals (CIs) for each parameter (Supplementary Tables S3-S8).
Estimates were considered significant when their 95\% CIs did not overlap with zero.

@@

In CR-DBH allometry, angiosperms exhibited a positive relationship between the scaling factor $a$ and wood density and a negative association between the exponent parameter  $b$ and wood density (@fig-wood1 g, h; @fig-wood2 b).
In gymnosperms, we found a positive association between wood density and the exponent parameter $b$, whereas neither the asymptote parameter $a$ nor the scaling constant $k$ showed a significant relationship with this functional trait (@fig-wood1 i, j, k; @fig-wood2 c).

@@

AGB estimates based on the Weibull model ($\text{AGB}_\text{H-WB}$) showed the highest predictive accuracy based on both the lowest root mean square error (RMSE) of `r sprintf("%.3g", data_metrics$metrics_agb_ang$rmse_wb_h)` Mg and the smallest bias of `r sprintf("%.3g", data_metrics$metrics_agb_ang$bias_wb_h * 100)`% (@fig-agb a).
Estimates based on the power-law model ($\text{AGB}_\text{H-PL}$) also showed competitive performance but had higher RMSE and bias compared to the Weibull-based approach, with RMSE of `r sprintf("%.3g", data_metrics$metrics_agb_ang$rmse_pl_h)` Mg and bias of `r sprintf("%.3g", data_metrics$metrics_agb_ang$bias_pl_h * 100)`% (@fig-agb b).
In contrast, models that used estimated DBH performed substantially worse (@fig-agb c-f).

@@

Our results supported a saturating H-DBH relationship in both angiosperm and gymnosperm species, consistent with theoretical expectations [@Falster2003; @Niklas2007] and empirical observations [e.g., @Feldpausch2011; @MartinezCano2019; @Song2024].
A recent study using Tallo suggested the Weibull function as the best model for six plant functional types (PFTs) [@Song2024].
While PFTs-based approaches can illuminate large-scale patterns, they may obscure species-specific variation in allometric scaling, and locally abundant species may bias the overall estimates.
Our best predictive model effectively captures species-specific patterns (Supplementary Table S6) despite underestimating the height of individuals with DBH > 30 cm due to substantial interspecific variation in H–DBH allometric scaling (@fig-compare).
This highlights the advantages of a multilevel modeling approach in addressing systematic differences across species.
Additionally, the metabolic scaling theory (MST) posits logarithmic scaling relationships among different axes of tree size, with an exponent close to 2/3 ($H \propto DBH^{2/3}$, $CR \propto DBH^{2/3}$), based on assumptions regarding mechanical and hydraulic constraints to plant growth [@West1999].
However, our findings revealed that the power-law model exponent of angiosperms (*b* = 0.538) was significantly smaller, while the actual exponent in gymnosperms (*b* = 0.704) was larger than the theoretical value of 2/3 (Supplementary Table S3).
This divergence highlights the importance of separately modeling the allometric relationships of each functional group to accurately capture their distinct scaling patterns.
Moreover, the power-law function performed worse than the saturating functions (Table 1), indicating that the actual H-DBH relationship does not follow the simple scaling laws proposed by MST.

Our analysis also revealed divergent patterns in CR-DBH allometry, highlighting fundamental differences in growth strategies and ecological adaptations between angiosperms and gymnosperms.
In angiosperms, the CR–DBH relationship follows a power-law form, suggesting a capacity for plastic crown expansion in competitive environments [e.g., @Poorter2012a; @Lines2012; @Hulshof2015], a conclusion consistent with a study in tropical moist forests by @MartinezCano2019.
In addition, the estimated exponent aligns closely with MST predictions for crown radius, supporting the selection of the power-law model for angiosperms, where the community-level exponent *b* is near the expected 2/3
($b_{\text{ang}}$ = `r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Median)`
[`r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Lower)`,
`r sprintf("%.3f", data_posterior$'Crown Radius'$Angiosperm$'Power-Law'$Without$b$Intercept_CI$Upper)`], Table 2).
In contrast, the CR-DBH allometry in gymnosperms showed an asymptotic relationship best described by the gMM function, with the Weibull function also showing competitive predictive performance (Table 1).
This asymptotic pattern likely arises from structural constraints in gymnosperms' growth forms that favor vertical over lateral expansion [e.g., @Lines2012; @Hulshof2015; @Jucker2024], an adaptation that can mitigate risks such as branch breakage under heavy snow loads [e.g., @Mikola1938; @Loehle2016; @Tavankar2019].

@@

Our analysis indicated distinct patterns in which wood density has a greater influence on interspecific variation in angiosperms than in gymnosperms.
Variation in wood density has consequential implications for the mechanical properties of wood and hence influences the mechanical stability of stems and branches [@Sterck1998], ultimately determining the architecture of trees [e.g., @Poorter2006].
Angiosperms have denser wood and a wider wood density range than gymnosperms (see Supplementary Table S2), resulting in greater and more flexible mechanical support [e.g., @Gelder2006; @Swenson2007; @Mo2024].
This allows them to develop a more diverse height and crown to optimize light capture, particularly in environments where light competition is intense, such as tropical regions.
In contrast, gymnosperms predominantly live and often dominate in extreme environments such as boreal and high-elevation regions [@Swenson2007; @Mo2024], prioritising survival and longevity [@Fossdal2024].
Due to their simpler and more homogeneous wood anatomy, these species display lighter wood and a narrower range of wood density [e.g., @Mo2024], which may result in less flexibility in the crown architecture.
These differences underscore how the contrasting evolutionary strategies of angiosperms and gymnosperms shape their architectural diversity, reflecting adaptations to distinct ecological niches.
